% Version 2019-01-08
% update – 161114 by Ken Arroyo Ohori: made spacing closer to Word template throughout, put proper quotes everywhere, removed spacing that could cause labels to be wrong, added non-breaking and inter-sentence spacing where applicable, removed explicit newlines
% update – 010819 by Dennis Wittich: made spacing and font size closer to Word template, updated references and refernces style
% update – 042319 by Dennis Wittich: font size of captions set to 'small', first author names are shortened, hyphenation fixed

\documentclass{isprs} % isprs class modified 23-04-2019 (Dennis Wittich)
\usepackage{subfigure}
\usepackage{setspace}
\usepackage{geometry} % added 27-02-2014 Markus Englich
\usepackage{epstopdf}
\usepackage[labelsep=period]{caption}  % added 14-04-2016 Markus Englich - Recommendation by Sebastian Brocks
\usepackage[british]{babel} 

\geometry{a4paper, top=25mm, left=20mm, right=20mm, bottom=25mm, headsep=10mm, footskip=12mm} % added 27-02-2014 Markus Englich
%\usepackage{enumitem}

%\usepackage{isprs}
%\usepackage[perpage,para,symbol*]{footmisc}

%\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\captionsetup{justification=centering,font=normal} % thanks to Niclas Borlin 05-05-2016
\captionsetup[figure]{font=small} % added 23-04-2019 Dennis Wittich
\captionsetup[table]{font=small} % added 23-04-2019 Dennis Wittich

\begin{document}

\title{SMODERP2D SOIL EROSION MODEL ENTERING AN OPEN SOURCE ERA WITH
  GPU-BASED PARALLELIZATION}

% KAO: Remove extra spacing
\author{
 M. Landa\textsuperscript{1}, J. Jeřábek\textsuperscript{2}, O. Pešek\textsuperscript{1}, P. Kavka\textsuperscript{2}}

% KAO: Remove extra newline
\address{
  \textsuperscript{1 }Dept.\ of Geomatics, Faculty of Civil Engineering, Czech Technical University in Prague,\\ Czech Republic - (martin.landa, ondrej.pesek)@fsv.cvut.cz\\
  \textsuperscript{2 }Dept.\ of Landscape Water Conservation, Faculty of Civil Engineering, Czech Technical University in Prague,\\ Czech Republic - (jakub.jerabek, petr.kavka)@fsv.cvut.cz\\
}

% If the corresponding author is NOT the final author, always add a % space before the subsequent comma, i.e.
% first author name\textsuperscript{a,}\thanks{Corresponding author} , % second author name \textsuperscript{b}, etc.
% thanks to Niclas Borlin 05-05-2016


\commission{VI, }{VI} %This field is optional.
\workinggroup{VI/4} %This field is optional.
\icwg{}   %This field is optional.

% KAO: Use times symbol
\abstract{ SMODERP2D is a runoff-soil erosion physically-based
  distributed episodic model used for calculation and prediction
  processes at agricultural areas and small watersheds. The core of
  the model is a raster based cell-by-cell mass balance calculation
  which includes the key hydrological processes, such as effective
  precipitation, surface runoff and stream network routing. Effective
  precipitation, the forcing of the runoff and erosion processes, is
  reduced by surface retention and infiltration. Surface runoff
  consists of two components: slower sheet and concentrated rapid rill
  flow. Stream network routing is performed line-by-line in user
  predefined polyline layer.

  SMODERP is a long-term running project driven by the Department of
  Landscape Water Conservation at the Czech Technical University in
  Prague. At the beginning SMODERP has been developed as a surface
  runoff simulated by profile model (1D). Later the model has been
  redesigned using spatially distributed method. This version is named
  SMODERP2D. Ongoing development is focused on obtaining parameters of
  the hydrological models, incorporating new infiltration and flow
  routing routines, and conceptualization of a rill flow and rill
  development. The model belongs to a family of so called GIS-based
  hydrological models utilizing capabilities of GIS software for
  geodata processing. Importantly, the SMODERP2D project is currently
  entering the open source world. Originally the model could be run
  only in proprietary Esri ArcGIS platform. A new version of the model
  presented by this contribution adds support for two key open source
  GIS platforms, GRASS GIS and QGIS. A newly developed GRASS module
  and QGIS plugin significantly increases accessibility of the
  SMODERP2D model for research purposes and also for engineering
  practice.

  Middle scale distributed hydrological models often encounter with a
  high computation costs and long model runtime. Long runtime is
  caused by high resolution input data which is easily available
  nowadays. The project also includes an experimental version of the
  SMODERP2D model enabling the parallelization of computations. This
  parallelization is done using TensorFlow, and its goal is to
  decrease the time needed for its run. It is supported by both CPU
  and GPU. Parallelization of computations is an important step
  towards providing SMODERP2D web processing services in order to
  allow quick and easy integration to highly specialized platforms
  such as Atlas Ltd.  }

\keywords{Hydrology, Soil Erosion Models, GIS, Open Source, Parallel computing}

\maketitle

%\saythanks % added 28-02-2014 Markus Englich

\section{INTRODUCTION}\label{INTRODUCTION}
 
% KAO: Sloppy spacing ensures non-overfull lines. Can be removed if this is not an issue.
\sloppy

\input{desc_introduction}

\section{MATERIAL AND METHODS}\label{sec:mat_met}

\input{desc_smoderp}

\input{desc_open_source}

\input{desc_parallel_expr}

\section{CONCLUSION}

\input{desc_conclusion}

\section*{ACKNOWLEDGEMENTS}\label{ACKNOWLEDGEMENTS}
The research has been supported by the research grants TJ01000270,
QK1910029, and internal CTU grant SGS17/173/OHK1/T3/11.

{
  \begin{spacing}{1.17}
    \normalsize
    \bibliography{smoderp2d_foss4g_2019} % Include your own bibliography (*.bib), style is given in isprs.cls
  \end{spacing}
}

\vspace{1cm}
\textit{Revised April 2019}

\end{document}
